---
title: "Filter data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filter data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, results='hide'}
library(camtrapviz)
library(ggiraph)
library(dplyr)
library(ggplot2)
```

This vignette demonstrates how to filter camera trap data based on 4 criteria:

-   cameras
-   species
-   observation type
-   date

## Import data

```{r}
data(mica, package = "camtraptor")
```

Here is the initial data:

```{r}
p <- plot_points(mica$data$observations,
                 camera_col = "deploymentID",
                 timestamp_col = "timestamp",
                 tooltip_info = "vernacularNames.en",
                 interactive = TRUE)
girafe(ggobj = p)
```

## Filter cameras

We can filter out cameras (NB: it is also possible to give a single camera in input).

Below are the cameras list for the original files:

```{r}
# Original cameras in observations and 
unique(mica$data$observations$deploymentID)

# Original cameras in deployments 
unique(mica$data$deployments$deploymentID)
```

```{r}
cam_filter <- filter_data(mica, 
                          cam_col_rec = "deploymentID", 
                          cam_filter = c("29b7d356-4bb4-4ec4-b792-2af5cc32efa8", "577b543a-2cf1-4b23-b6d2-cda7e2eac372"))
```

The cameras have been filtered out from both observations and deployments tables:

```{r}
unique(cam_filter$data$observations$deploymentID)

unique(cam_filter$data$deployments$deploymentID)
```

```{r}
plot_points(cam_filter$data$observations,
            camera_col = "deploymentID",
            timestamp_col = "timestamp")
```

## Filter species

We can filter out by species (removing one or several species):

Below is the initial species list:
```{r}
unique(mica$data$observations$vernacularNames.en)
```


```{r}
spp_filter <- filter_data(mica, 
                          spp_col = "vernacularNames.en", 
                          spp_filter = "great herons")
```

After data cleaning, `great herons` was removed.
```{r}
unique(spp_filter$data$observations$vernacularNames.en) 
```


## Filter observation type

We can filter out by observation type. It can be useful for instance if we want to keep only animals (and not unidentifiable, humans...).

Below are the initial values of observation types:

```{r}
obstype <- mica$data$observations$observationType
unique(obstype)
```

Let's filter out all categories except `animal`:

```{r}
# Define the types we want to remove
obs_out <- unique(obstype[obstype != "animal"])
obs_out

obs_filter <- filter_data(mica, 
                          obs_col = "observationType", 
                          obs_filter = obs_out)
```

The only observation type remaining after data cleaning is `animal`.

```{r}
unique(obs_filter$data$observations$observationType)
```

## Filter dates

It is also possible to specify date bounds to the data.

Here is the original date range:

```{r}
# Original date range
(old_range <- range(mica$data$observations$timestamp))
```


Let's filter a subset of the data between January, 1st 2020 and January, 1st 2021.

```{r}
new_range <- c("2020-01-01", "2021-01-01")
```


We visualize this filter on a plot: solid lines are the old data bounds and dashed lines represent the new date range.

```{r}
plot_points(mica$data$observations, 
            camera_col = "deploymentID",
            timestamp_col = "timestamp") +
  ggplot2::geom_vline(xintercept = as.POSIXct(new_range, 
                                              tz = "UTC"),
                      linetype = "dashed") +
  ggplot2::geom_vline(xintercept = old_range)
```


Let's filter the data.

```{r}
date_filter <- filter_data(mica, 
                           timestamp_col = "timestamp", 
                           daterange = new_range)
```

The `daterange` argument can also be a `Date` or `POSIX`:

```{r}
(new_range_date <- as.Date(new_range))
date_filter <- filter_data(mica, 
                           timestamp_col = "timestamp", 
                           daterange = new_range_date)

(new_range_posix <- as.POSIXct(new_range))
date_filter <- filter_data(mica, 
                           timestamp_col = "timestamp", 
                           daterange = new_range_posix)
```


The new data range is comprised within the specified bounds (January, 1st 2020 and January, 1st 2021).

```{r}
# The new range is restricted
range(date_filter$data$observations$timestamp)
```

It is also possible to filter data that has date and times i two columns. First, we create a dataset with these two columns:

```{r}
mica_date_time <- mica

mica_date_time$data$observations <- mica_date_time$data$observations |> 
  dplyr::mutate(date = format(timestamp, "%F"),
                time = format(timestamp, "%T"),
                .after = "timestamp")

mica_date_time$data$observations |> 
  dplyr::select(timestamp, date, time) |> 
  head()
```

And we filter dates on these data.

```{r}
date_filter <- filter_data(mica_date_time, 
                           date_col = "date",
                           time_col = "time", 
                           daterange = new_range)
```


The new data range is the same as above.

```{r}
range(date_filter$data$observations$timestamp)
```


## Combine filters

Finally, it is possible to combine multiple conditions to customize data filtering:

```{r}
custom_filter <- filter_data(mica,
                             cam_filter = "29b7d356-4bb4-4ec4-b792-2af5cc32efa8",
                             cam_col_rec = "deploymentID",
                             spp_filter = c("great herons", "grey heron"),
                             spp_col = "vernacularNames.en",
                             timestamp_col = "timestamp",
                             daterange = c("2020-01-01", "2021-04-18"))
```


Let's check the filtering:

```{r}
# Full dataset
unique(mica$data$observations$deploymentID)
# Filtered dataset
unique(custom_filter$data$observations$deploymentID)

# Full dataset
unique(mica$data$observations$vernacularNames.en)
# Filtered dataset
unique(custom_filter$data$observations$vernacularNames.en)

# Full dataset
range(mica$data$observations$timestamp)
# Filtered dataset
range(custom_filter$data$observations$timestamp)
```

