---
title: "Read and clean data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read and clean data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates how the functions included in this package can be used to read and clean different data formats.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(camtrapviz)
```

## Write data in tempfile

```{r}
# csv files ------------------------------------------
data(recordTableSample, package = "camtrapR")
data(camtraps, package = "camtrapR")

# Create subfolder
dir.create(paste0(tempdir(), "/csv"))

# Write files
recordfile <- paste0(tempdir(), "/csv/records.csv")
camtrapfile <- paste0(tempdir(), "/csv/camtraps.csv")

write.csv(recordTableSample, recordfile, 
          row.names = FALSE)
write.csv(camtraps, camtrapfile, 
          row.names = FALSE)
```


```{r}
# csv camera ------------------------------------------
# Create file
recordcam <- recordTableSample |>
  dplyr::left_join(camtraps, by = "Station")

# Create subfolder
dir.create(paste0(tempdir(), "/csvcam"))

# Write file
recordcamfile <- paste0(tempdir(), "/csvcam/recordcam.csv")
write.csv(recordcam, recordcamfile, 
          row.names = FALSE)
```


## Records and cameras in separate csv files (2 csv files)

### Read data

```{r}
dat <- read_data(records_path = recordfile,
                 cameras_path = camtrapfile,
                 sep_records = ",", sep_cameras = ",")
```

```{r}
head(dat$data$observations)
head(dat$data$deployments)
```

The imported file is a list with one component `$data` containing 2 dataframes:

+ `$observations` contains the records
+ `$deployments` contains the cameras information

### Clean data

This step ensures all columns have the desired type. It will also move these columns to the beginnning of the table.

```{r}
rec_type <- list(Station = "as.character",
                 Date = list("as_date",
                             format = "%Y-%m-%d"),
                 Time = "times",
                 DateTimeOriginal = "as.POSIXct")

cam_type <- list(Station = "as.character",
                 Setup_date = list("as.Date",
                                   format = "%d/%m/%Y"), 
                 Retrieval_date = list("as.Date",
                                       format = "%d/%m/%Y"))

dat_clean <- clean_data(dat, 
                        rec_type = rec_type,
                        cam_type = cam_type)
```

```{r}
head(dat_clean$data$observations)
head(dat_clean$data$deployments)
```

## Records and cameras in the same csv (1 csv file)

### Read data

```{r}
dat <- read_data(records_path = recordcamfile,
                 sep_records = ",")
```

```{r}
head(dat$data$observations)
head(dat$data$deployments)
```

Again the imported file is a list with one component `$data`:

+ `$data$observations` contains the cameras and records information
+ `$data$deployments` is `NULL`

### Clean data

Here this step will split the information between cameras and records. It will move all columns listed in `cam_type` in the cameras file.

```{r}
cam_type2 <- list(Station = "as.character",
                  Setup_date = list("as.Date",
                                    format = "%d/%m/%Y"), 
                  Retrieval_date = list("as.Date",
                                        format = "%d/%m/%Y"),
                  utm_y = "as.numeric",
                  utm_x = "as.numeric",
                  Problem1_from = list("as.Date",
                                       format = "%d/%m/%Y"),
                  Problem1_to = list("as.Date",
                                     format = "%d/%m/%Y"))

dat_clean <- clean_data(dat, 
                        rec_type = rec_type,
                        cam_type = cam_type2,
                        cam_col_records = "Station",
                        split = TRUE)
```

```{r}
head(dat_clean$data$observations)
head(dat_clean$data$deployments)
```

## CamtrapDP format (json file)

### Read data
```{r}
dat <- read_data(records_path = "https://raw.githubusercontent.com/tdwg/camtrap-dp/main/example/datapackage.json")
```

Internally, we use the function `read_camtrap_dp` from the `camtraptor` package. The imported object is a `list` and the observations and deployments info are in the `$data` slot.

```{r}
class(dat)

head(dat$data$observations)
head(dat$data$deployments)
```

### Clean data

Here, the data follows the camtrapDP standard and does not need cleaning. However, for this demonstration we will change the time stamp type:

```{r}
dat$data$observations$timestamp <- as.character(dat$data$observations$timestamp)

# timestamp was converted to character
class(dat$data$observations$timestamp)
```


```{r}
rec_type <- list(timestamp = "as.POSIXct")

dat_clean <- clean_data(dat, 
                        rec_type = rec_type)

# timestamp is back to POSIX
class(dat_clean$data$observations$timestamp)
```

```{r}
head(dat_clean$data$observations)
head(dat_clean$data$deployments)
```