---
title: "Camera trap data analysis"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
---

<style type="text/css">
  body{
    font-family: 'Ubuntu', 'Helvetica Neue', Helvetica, Arial, sans-serif;
}

  .list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #006d77 !important;
    border-color: #006d77 !important;
  }
  
  .btn {
    font-family: inherit !important;
    color: #adadad !important;
    background-color: #f5f5f5 !important;
    border-color: #ccc !important;
    padding: 2px 3px 2px 3px !important;
    font-size: 13px !important;
  }
  
  .btn:hover {
    color: #fff !important;
    background-color: #798d8f !important;
    border-color: #798d8f !important;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  out.width = "100%", 
  tidy = TRUE,
  class.source='fold-show'
)
has_lonlat <- {{has_lonlat}}
```

```{r libraries, message=FALSE, results='hide'}
library(camtrapviz)
library(ggiraph)
library(ggplot2)
library(dplyr)
```

## Data import

```{r data_reading}
{{data_reading}}
```

```{r data_cleaning}
{{data_cleaning}}
```

## Data filtering

```{r data_filtering}
{{data_filtering}}
```

Below is a plot with the filtered out data in grey and the kept data in black.

```{r filtering_plot, class.source = NULL}
{{filtering_plot}}
```

## Data summary

### Some statistics

Some statistics on the project: 

+ Number of cameras: {{ncameras}}
+ Number of species: {{nspecies}}
+ Trapping nights: {{sampling_length}}
+ Active from {{daterange}}

### Summary  tables

Here is a summary table per camera:

```{r cameras_table}
{{camtable}}
```

Here is a summary table per species:

```{r species_table}
{{spptable}}
```

`r if(has_lonlat){"### Map\nThis map shows the cameras of this project that have spatial coordinates."}`

```{r plot_map_all, class.source = NULL, eval = has_lonlat}
{{plot_map_all}}
```

### Species occurrences plot

This plot shows the occurrences of species on the different cameras.

```{r plot_occurrences, class.source = NULL}
{{plot_occurrences}}
```

## Community analyses

### Species {{count_unit}} count
This graph shows the {{count_unit}} count for each species.

```{r, class.source = NULL}
{{plot_species_bars}}
```

### Diversity analyses

The table below summarizes diversity indices for each camera.

```{r}
{{diversity_table}}
```

{{diversity_index}} is shown on the plot below:

```{r, class.source = NULL}
{{plot_diversity}}
```

## Unispecific analyses

Now the analyses will concentrate on {{focus_spp}}. We filter the subset of the data for this species:

```{r}
{{focus_spp_records[[1]]}}
```

### Activity

Below is the activity curve of {{focus_spp}}, estimated using a von Mises kernel probability density, along with the histogram of the observed activity times.

```{r, class.source = NULL}
{{density_plot[[1]]}}
```

### Abundance per camera

{{abundance_value}} of {{focus_spp}} at each camera is summarized in the plot below:

```{r, class.source = NULL}
{{abundance_plot[[1]]}}
```
