---
title: "Camera trap data analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  out.width = "100%", 
  tidy = TRUE
)
```

```{r libraries}
library(camtrapviz)
library(ggiraph)
```

## Data import

```{r data_cleaning}
{{data_cleaning}}
```

## Data filtering
```{r data_filtering}
{{data_filtering}}
```

## Data summary

Here is a summary table per camera:

```{r cameras_table}
{{camtable}}
```

Here is a summary table per species:

```{r species_table}
{{spptable}}
```

### Map

This map shows the cameras of this project that have spatial coordinates.

```{r plot_map, echo = FALSE}
{{plot_map}}
```

### Species occurrences

This plot shows the occurrences of species on the different cameras.

```{r plot_occurrences, echo = FALSE}
{{plot_occurrences}}
```

## Community analyses

This graph shows the how many individuals of each species were seen.

```{r, echo = FALSE}
{{plot_species_bars}}
```


The table below summarizes diversity indices for each camera.

```{r}
{{diversity_table}}
```

Below is the plot representing the chosen index:

```{r, echo = FALSE}
{{plot_diversity}}
```

